---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# cplselection

<!-- badges: start -->
<!-- badges: end -->

**cplselection** is an R package that corrects sample selection bias in regression estimates using a Bayesian Gaussian copula method. Different from the Heckit correction method, this method allows user to flexibly specify the selection and the outcome marginal distributions deviating from the bivariate normality assumption. 

Available options for selection distribution are 

- binomial distribution (with probit link, logit link, or complementary log-log link)

Available options for outcome distribution are

- binomial distribution (with probit link, logit link, or complementary log-log link)
- Poisson distribution
- negative binomial distribution
- exponential distribution 
- normal distribution


## Installation

To install the **cplselection** package from GitHub, you need to have R and RStudio installed on your system. Follow these steps to install the package:

### Prerequisites

Make sure you have the `devtools` or `remotes` package installed. You can install it from CRAN if you haven't already:

```r
install.packages("devtools")  # or use remotes: install.packages("remotes")
``` 

### Install and Load the Package

Now you are ready to install the **cplselection** package:

``` r
# Using devtools
devtools::install_github("codingcindy/cplselection")

# Alternatively, using remotes
remotes::install_github("codingcindy/cplselection")
```

After installation, you can load the package using:

```r
library(cplselection)
```

## Usage

The **cplselection** package provides two functions for the estimation procedure. The function `cplselectionMCMC` performs MCMC iterations and returns an R list object containing the MCMC draws and information on acceptance probabilities and step sizes. The other function `cplselectionInfer` makes inferences based on the MCMC draws returned by `cplselectionMCMC` function and returns the Bayesian estimates in an R data.frame object.

Here’s a quick overview of how to use these functions: 

### Example Usage

1. Generate example data set
  ```{r example}
  # Example data : Poisson outcome variable YO, logit selection process YS
  library(cplselection)
  nobs <- 300
  dependence <- 0.5
  betaO0 <- 1
  betaO1 <- 1
  betaS0 <- 0
  betaS1 <- 1
  betaS2 <- 1
  uo <- runif(nobs)
  us <- pnorm(dependence*qnorm(uo) + dependence*rnorm(nobs))
  data <- data.frame(X1 = rnorm(nobs), X2 = rnorm(nobs))
  data$YO <- qpois(p=uo, lambda=exp(betaO0 + betaO1*data$X1))
  data$YS <- qbinom(p=us, prob=1/(1+exp(-(betaS0 + betaS1*data$X1 + betaS2*data$X2))), size=1)
  data$YO[data$YS==0] <- NA
  summary(data)
  ```
2. Perform MCMC draws with `cplselectionMCMC`
  ```{r}
  # perform MCMC draws
  loop <- 10000
  draws <- cplselectionMCMC(
    outcome_formula = YO~X1, select_formula = YS~X1+X2,
    outcome_dist = "Poisson", select_dist = "Logit", 
    data = data, loop = loop)
  ```
3. Bayesian inference with `cplselectionInfer`
  ```{r}
  # if the true values are not known
  result.1 <- cplselectionInfer(iterates=draws, burnin=loop/2)
  print(round(result.1, 2))
  
  # if the true values are known
  trueval <- c(dependence, betaO0, betaO1, betaS0, betaS1, betaS2)
  result.2 <- cplselectionInfer(iterates=draws, burnin=loop/2, trueval=trueval)
  print(round(result.2, 2))
  ```

## Documentation

More details can be found by running: 

  ```{r, eval=FALSE}
  ?cplselectionMCMC
  ?cplselectionInfer
  ```

## License

This package is licensed under GPL-3 (https://www.gnu.org/licenses/gpl-3.0.html).
